Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-09-29 apj.tex
100 simulation version

The significance of the $\vec{\delta}$ measurement is evaluated using simulated datasets.
For one set of tests, we generate 50 random realizations of data with input  parameters equal to the median values of the 
parameter posteriors found in
\S\ref{model:sec}, and using the same noise as our data.  
\color{red}
(Given the non-trivial computational time for the Monte Carlo, the 2\% precision from the 
is deemed sufficient for our purposes.)
\color{black}
To represent the null hypothesis, the input $\vec{\delta}$-vector is set to zero.
Figure~\ref{pvalues_delta2:fig} shows results of fits of these null-realizations next  to the results from the real data.
One statistic shown is the median and 68\% credible interval of the squared amplitude  $\delta^2$.
As $\delta^2$ measures the magnitude response due to the third parameter, it is expected to have a small value
when such a new parameter is not relevant.
Comparing separately the  $-1$, 0, and $1$-sigma
values of  the credible interval of our data relative to those from the simulations, we obtain
three different $p$-values ranging from 0.90--0.98, indicating that the
\color{red}
null hypothesis is a poor representation of the data.
\color{black}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.9in]{output25_sim_null/pvalues_delta2.pdf} 
  \includegraphics[width=2.9in]{output25_sim_null/pvalues_relperp.pdf} 
            \caption{ Left: The squared length, $\delta^2$;  and Right: the relative length perpendicular to the plane defined by $\vec{\gamma}$, $\delta_\perp^2/\delta^2$, for 50 independent simulated data sets with input $\delta=0$.  The fits
            are ordered by the median value, the null realizations are shown in blue and the result from our data is shown in red.
            \label{pvalues_delta2:fig}}
\end{figure}

An interesting feature of $\vec{\delta}$ is that its direction lies close to the plane defined by the
$\vec{\gamma}$ vectors.
This leads to the concern that our model and analysis  could spontaneously generate a third
vector almost degenerate with the others, despite the prior 
\color{red}
encouraging
\color{black}
 the independence of $k_0$, $k_1$, and $D$.
This \color{red}
possibility
\color{black}
 is tested using the 
relative projection of $\vec{\delta}$  perpendicular to the plane defined by the $\vec{\gamma}$'s, $\delta_\perp^2/\delta^2$,
for the same 50 null simulations that were just described.
Figure~\ref{pvalues_delta2:fig} shows  that for none of the simulations does our analysis produce 
 $\delta_\perp^2/\delta^2$ as low as that from our data.  The $p$-values $>0.98$ indicate that 
 \color{red}
 the close alignment of $\vec{\delta}$  with  $\vec{\gamma}$'s is unlikely to be by chance. 
 \color{black}

Potential biases in our results are measured through simulation.
We generate 50 random realizations of data with input  parameters equal to the median values found in
this Section, except for $\vec{\delta}$, which is assigned
\color{red}
the mode shown in Figure~\ref{deltacorner:fig},
\color{black}
 and the same noise as our data.  The posterior for each realization is analyzed individually, and the posteriors of all realizations are stacked.  The stack is sensitive to
biases in the best fit value.
The stack of  $\vec{\delta}$  is shown in Figure~\ref{simdelta:fig}.  There is a bias between
the input  $(-1.12, -0.89, -0.36,  -0.16, 0.21)$ and the mode of  the stacks $(-1.06, -0.95, -0.41, -0.11, 0.40)$.
We do not consider this bias to be an issue in our detection of a third parameter, as the difference is significantly
smaller than the statistical uncertainty of a single analysis and the length of the vector changes by $< 3\%$.
The posterior stacks for the other global parameters 
are shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties. 
Note that some bias is expected: the priors for the distributions of $\Delta$, $k$, and $D$ are inconsistent
with the extracted distributions.  Also, 
there is a tendency for the slope parameters associated with the spectral features to be slightly underestimated, which is an expected consequence of 
the uncertainties in the predictor variable that lead to 
regression dilution \citep{spearman04}.

>>>>

The significance of the $\vec{\delta}$ measurement is evaluated using simulated datasets.
For one set of tests, we generate 100 random realizations of data with input  parameters equal to the median values of the 
parameter posteriors found in
\S\ref{model:sec}, and using the same noise as our data.  
To represent the null hypothesis, the input $\vec{\delta}$-vector is set to zero.
Figure~\ref{pvalues_delta2:fig} shows results of fits of these null-realizations next  to the results from the real data.
One statistic shown is the median and 68\% credible interval of the squared amplitude  $\delta^2$.
As $\delta^2$ measures the magnitude response due to the third parameter, it is expected to have a small value
when such a new parameter is not relevant.
Comparing separately the  $-1$, 0, and $1$-sigma
values of  the credible interval of our data relative to those from the simulations, we obtain
three different $p$-values ranging from 0.88--0.98, indicating that the
\color{red}
null hypothesis is a poor representation of the data.
\color{black}

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.9in]{output25_sim_null/pvalues_delta2.pdf} 
  \includegraphics[width=2.9in]{output25_sim_null/pvalues_relperp.pdf} 
            \caption{ Left: The squared length, $\delta^2$;  and Right: the relative length perpendicular to the plane defined by $\vec{\gamma}$, $\delta_\perp^2/\delta^2$, for 100 independent simulated data sets with input $\delta=0$.  The fits
            are ordered by the median value, the null realizations are shown in blue and the result from our data is shown in red.
            \label{pvalues_delta2:fig}}
\end{figure}

An interesting feature of $\vec{\delta}$ is that its direction lies close to the plane defined by the
$\vec{\gamma}$ vectors.
This leads to the concern that our model and analysis  could spontaneously generate a third
vector almost degenerate with the others, despite the prior 
\color{red}
encouraging
\color{black}
 the independence of $k_0$, $k_1$, and $D$.
This \color{red}
possibility
\color{black}
 is tested using the 
relative projection of $\vec{\delta}$  perpendicular to the plane defined by the $\vec{\gamma}$'s, $\delta_\perp^2/\delta^2$,
for the same 100 null simulations that were just described.
Figure~\ref{pvalues_delta2:fig} shows  that for 
\color{red}
 $-1$, 0, and $1$-sigma
values of  the credible interval, only 0--2\% of the simulations have values of  $\delta_\perp^2/\delta^2$ as low as our data.
\color{black}
The $p$-values $>0.98$ indicate that 
 \color{red}
 the close alignment of $\vec{\delta}$  with  $\vec{\gamma}$'s is unlikely to be by chance. 
 \color{black}

Potential biases in our results are measured through simulation.
We generate 100 random realizations of data with input  parameters equal to the median values found in
this Section, except for $\vec{\delta}$, which is assigned
\color{red}
the mode shown in Figure~\ref{deltacorner:fig},
\color{black}
 and the same noise as our data.  The posterior for each realization is analyzed individually, and the posteriors of all realizations are stacked.  The stack is sensitive to
biases in the best fit value.
The stack of  $\vec{\delta}$  is shown in Figure~\ref{simdelta:fig}.  There is a bias between
the input  $(-1.12, -0.89, -0.36,  -0.16, 0.21)$ and the mode of  the stacks $(-1.2, -1.06, -0.49, -0.25, 0.21)$.
We do not consider this bias to be an issue in our detection of a third parameter, as the difference is significantly
smaller than the statistical uncertainty of a single analysis and the length of the vector changes by $< 3\%$.
The posterior stacks for the other global parameters 
are shown in Figures~\ref{simglobal1:fig}--\ref{simglobal5:fig}.  The input parameters are recovered with negligible bias
relative to the statistical uncertainties. 
Note that some bias is expected: the priors for the distributions of $\Delta$, $k$, and $D$ are inconsistent
with the extracted distributions.  Also, 
there is a tendency for the slope parameters associated with the spectral features to be slightly underestimated, which is an expected consequence of 
the uncertainties in the predictor variable that lead to 
regression dilution \citep{spearman04}.
